Embedding Intuitionistic into Classical Logic

### 2023-05-03 LPAR/LPAR-EIICL.tex
changes from reviews

\documentclass{easychair}
\usepackage{amsmath, amsfonts, amsthm}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage[T1]{fontenc}
\usepackage{todonotes}
%
\usepackage{tikz}
\usetikzlibrary{shapes,backgrounds}

@@

\begin{definition}
	A \emph{Kripke structure} $\mathcal K = (W, (v_w)_{w\in W})$ consists of a partially ordered set $W$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that $v_u(A)\leq v_w(A)$ for all $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
	\end{itemize}
	We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulae with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.

@@

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set $W$, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
	\begin{itemize}
		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/x]\models\varphi$.
	\end{itemize}
	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$. $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
	We denote the set of valid formulae with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to propositional tautologies there are now quantified formulae which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.

@@

Finally we recall a result from~\cite{otten2005clausal}, on which we build in this paper.

@@

\begin{theorem}
	\label{thm:reduction-first-order-short}
	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and $\varphi$ contains a new binary predicate $E$ as well as a number of new function symbols. Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
\end{theorem}

@@

\begin{theorem}\label{thm:prop-translation}
	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
	with flat $\mathcal R$ and implicational $\mathcal S$. Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$. For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $A^\lambda$. Define $\mathcal R^\#$ as the union over all sets
	$$
		\left\{\bigwedge_iA_i^\lambda\to\bigvee_jB_j^\lambda\:|\:\lambda\in\Lambda\right\}
	$$
	for $\bigwedge_iA_i\to\bigvee_jB_j\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
	$$
		\left\{\left(A^{\lambda\psi} \to B^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $\mathcal P^\#$ as the union over all sets
	$$
		\left\{A^\lambda\to B^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	 for each propositional variable $A$ occurring in $\varphi$, encoding persistency. Then
	$$
		\varphi^\# := \left(\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge \mathcal P^\#\right)\to P^\epsilon
	$$
	is classically valid if and only if $\varphi$ is intuitionistically valid.
\end{theorem}
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

@@

It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}, but we make it explicit. In particular, our translation enables a direct translation of counter-models. For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
The translation from the last section will serve as starting point for this section.

@@

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\psi^u$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is a new bound variable.
		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
		\item $(\forall x\varphi)^b = \forall w(i\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) := \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:
	\begin{align*}
		\text{PartialOrder}(\preceq) := &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) := &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) := &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) := &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) := &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	\[\varphi^C:= K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b\] where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulae $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
where $\vec r, s$ are new constant symbols, each formula in $\mathcal R^{bS}$ is of the form
$$\forall w \forall \vec x(s\preceq w\to\vec E(x, w)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)),$$
each formula $\psi\in\mathcal X^{bS}$ is of the form
$$
	\resizebox{\hsize}{!}{$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to (w\preceq f_\psi(\vec x, w)\to A(\vec a, f_\psi(\vec x, w))\to B(\vec b, f_\psi(\vec x, w)))\to C(\vec c, w))$}
$$
and each formula $\psi\in\mathcal Q^{bS}$ is of the form
$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec B, w))$$
with $f_\psi$ being a new function symbol.

@@

Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
$$
	\bigwedge\{\forall u\forall x\forall\vec y(E(x, u)\to E(x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\}
$$
and Persistent$(u, w)$ with
$$
	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a predicate in $\varphi$}\}.
$$Putting all of this together yields the following translation:
\begin{theorem}\label{thm:fo-translation}
	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above. For each $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name. For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$. Consider a new binary predicate $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal R^\#$ by including
	$$
		\forall u\forall \vec x\left(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)\right)
	$$
 	for each flat clause
 	$
 		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
 	$
 	$\mathcal X^\#$ by including
 	$$
 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec x, u))\to B(\vec b, f_\psi(\vec x, u))\to C(\vec c, u)\right)
 	$$
 	for each implication clause
	$
		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
	$
	and $\mathcal Q^\#$ by including
	$$
		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec x, u))\right)\to B(\vec b, u)\right)
	$$
	for each implication clause
	$
		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
	$
	Then $\varphi$ is intuitionistically valid if and only if
	$$
		\varphi^\# := K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
	$$
	is classically valid, where $s, \vec r$ are new constants introduced by Herbrandization of free variables,
	\begin{align*}
		K(\varphi) := &\:\bigwedge\left\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\right\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a predicate in $\varphi$}\right\}$.}
	\end{align*}
	The size of the obtained formula is linear in the input. However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary predicate has been extended to a $n+1$-ary predicate.
\end{theorem}

@@

We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} were used. All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem. The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of instance.

@@

In general, our approach falls short of the state of the art. 
However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$ resp. $|\mathcal X| \geq 6$.

>>>>

\documentclass{easychair}
\usepackage{amsmath, amsfonts, amsthm}
\usepackage{wrapfig}
\usepackage{multicol}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,backgrounds}

%used for d
\usepackage{todonotes}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\newcommand{\red}[1]{{\color{red} #1}}

@@

\begin{definition}
	A \emph{Kripke structure}  \sout{$\mathcal K = (W, (v_w)_{w\in W})$} \red{$\mathcal K = (\mathbf W, (v_w)_{w\in W})$} consists of a partially ordered set $\mathbf W = (W, \leq)$ of \emph{worlds}, called the Kripke \emph{frame}, and an indexed family of valuations $(v_w)_{w\in W}$ such that \sout{$v_u(A)\leq v_w(A)$} \red{$v_u(A) = 1$ entails $v_w(A) = 1$} for all worlds $u\leq w$ and propositional variables $A$ (this is called the \emph{persistency condition}).
	The model relation between the $v_u$ and formulae $\varphi$ is defined as before, except in the case of implications, where we set
	\begin{itemize}
		\item $v_u\models\varphi\to \psi$ iff for all $w\geq u$ we have $w\not\models\varphi$ or $w\models\psi$.
	\end{itemize}
	We say that $\varphi$ is satisfied at a world $u$ if $v_u\models\varphi$ and write $u\models\varphi$. If a formula is satisfied at every world then $\mathcal K$ is a \emph{model} for $\varphi$. A formula is valid if every Kripke structure is a model of it. We denote the set of valid formulae with \emph{IPC} (Intuitionistic Propositional Calculus).
\end{definition}
There are many classical tautologies which are not intuitionistically valid, e.g. the law of excluded middle $A\vee\neg A$. On the other hand any intuitionistic theorem is a classical one.

@@

\begin{definition}
	A $\Sigma$-\emph{Kripke structure} $\mathcal{K}$ is a partially ordered set \sout{$W$} \red{$(W, \leq)$}, called the Kripke \emph{frame}, and an indexed family of $\Sigma$-structures $(\mathcal{M}_w)_{w\in W}$ such that for $u\leq w$ we have $M_u\subseteq M_w$, $f^{I_w}|_{M_u} = f^{I_u}$ and $R^{I_w}|_{M_u} = R^{I_u}$.\footnote{Here $f|_M$ denotes the restriction of $f$ to $M$}
	The model relation between $\mathcal M_u$, $v$ (mapping into $M_u$) and formulae $\varphi$ is defined as before, except in the cases of implication and $\forall$-quantification, where
	\begin{itemize}
		\item $\mathcal M_u, v\models \varphi\to\psi$ iff for every $w\geq u$ we have $\mathcal M_w, v\not\models\varphi$ or $\mathcal M_w, v\models\psi$.
		\item $\mathcal M_u, v\models\forall x\varphi$ iff for every $w\geq u$ and $m\in M_w$ we have $\mathcal M_w, v[m/x]\models\varphi$.
	\end{itemize}
	We write $u, v\models\varphi$ if $\mathcal M_u, v\models \varphi$. $\mathcal{K}$ satisfies $\varphi$ if $u, v\models\varphi$ holds for every world $u\in W$ and variable assignment $v$. $\varphi$ is valid if it is satisfied by every Kripke structure.
	We denote the set of valid formulae with \emph{IQC} (Intuitionistic Quantified Calculus).
\end{definition}
In addition to propositional tautologies there are now quantified formulae which are classically valid but not intuitionistically, e.g., $\neg\forall x A(x)\to \exists x \neg A(x)$.

@@

Finally we recall a result from~\cite{claessen2015sat}, on which we build in this paper.

@@

\begin{theorem}
	\label{thm:reduction-first-order-short}
	There exists a linear-time procedure that gives for every first-order formula $\varphi$ a formula $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. The size of $\varphi^\#$ is in linear in the size of $\varphi$, however, for each $n$-ary predicate symbol in $\varphi$ there is a corresponding $n+1$-ary predicate symbol in $\varphi^\#$, and \sout{$\varphi$}\red{$\varphi^\#$} contains a new binary predicate $E$ as well as a number of new function symbols. Furthermore, there is an effective translation between intuitionistic counter-models of $\varphi$ and classical counter-models of $\varphi^\#$.
\end{theorem}

@@

\begin{theorem}\label{thm:prop-translation}
	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
	with flat $\mathcal R$ and implicational $\mathcal S$. Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$. For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional \sout{Variable}\red{variable} $A^\lambda$. Define $\mathcal R^\#$ as the union over all sets
	$$
		\left\{\bigwedge_iA_i^\lambda\to\bigvee_jB_j^\lambda\:|\:\lambda\in\Lambda\right\}
	$$
	for $\bigwedge_iA_i\to\bigvee_jB_j\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
	$$
		\left\{\left(A^{\lambda\psi} \to B^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $\mathcal P^\#$ as the union over all sets
	$$
		\left\{A^\lambda\to B^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	 for each propositional variable $A$ occurring in $\varphi$, encoding persistency. Then
	$$
		\varphi^\# := \left(\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge \mathcal P^\#\right)\to P^\epsilon
	$$
	is classically valid if and only if $\varphi$ is intuitionistically valid.
\end{theorem}
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

@@

It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}, but we make it explicit. In particular, our translation enables a direct translation of counter-models. For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
The translation from the last section will serve as \red{a} starting point for this section.

@@

\begin{definition}\label{def:fo-translation}
	Let $\varphi$ be a predicate formula. Define $\varphi^{b}$ inductively:
	\begin{itemize}
		\item $\bot^b = \bot$.
		\item $A(\vec t)^{b} = A(\vec t, b)$ for every predicate symbol $A$.
		\item $(\varphi\circ\psi)^b = \varphi^b\circ\sout{\psi^u}\red{\psi^b}$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^b = \forall w(b\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is a new bound variable.
		\item $(\exists x\varphi)^b = \exists x(E(x,b)\wedge\varphi^b)$.
		\item $(\forall x\varphi)^b = \forall w(\sout{i}\red{b}\preceq w\to \forall x(E(x, w)\to \varphi^w))$ where $w$ is a new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures as well as the predicate $E$, i.e.
	\begin{align*}
		K(\varphi) := \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
		& \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
	\end{align*}
	with e.g.:
	\begin{align*}
		\text{PartialOrder}(\preceq) := &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{DomainSubset}(u, w) := &\:\forall x(E(x, u)\to E(x, w))\\
		\text{World}(u) := &\:\exists xE(x, u)\\
		\text{DomainClosed}(u) := &\:\bigwedge\left\{\forall\vec x(\vec E(\vec x, u)\to E(f(\vec x), u))\:|\:\text{$f$ a function in $\varphi$}\right\}\\
		\text{Persistent}(u, w) := &\:\bigwedge\left\{\forall\vec x(\vec A(\vec x, u)\to A(\vec x, w))\:|\:\text{$A$ a predicate in $\varphi$}\right\}
	\end{align*}
	Then define
	\[\varphi^C:= K(\varphi)\to\text{World}(b)\to \vec E(\vec a, b)\to \varphi^b\] where $\vec a$ contains all free variables occurring in $\varphi$ and $b$ is a new free variable.
\end{definition}

@@

Now again due to Lemma~\ref{lemma:first-order-normal-form} it is sufficient to consider formulae $\varphi$ of the form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$ with flat clauses $\mathcal R$, implication clauses $\mathcal X$ and quantification clauses $\mathcal Q$, which after Herbrandization leaves us with $$\varphi^{CH} = K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge R^{bS}\wedge \bigwedge X^{bS}\wedge\bigwedge Q^{bS}\right)\to P(s)$$
where $\vec r, s$ are new constant symbols, each formula in $\mathcal R^{bS}$ is of the form
$$\forall w \forall \vec x(s\preceq w\to\vec E(x, w)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)),$$
each formula $\psi\in\mathcal X^{bS}$ is of the form
$$
	\resizebox{\hsize}{!}{$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to (w\preceq f_\psi(\vec x, w)\to A(\vec a, f_\psi(\vec x, w))\to B(\vec b, f_\psi(\vec x, w)))\to C(\vec c, w))$}
$$
and each formula $\psi\in\mathcal Q^{bS}$ is of the form
$$\forall w\forall\vec x(s\preceq w\to\vec E(\vec x, w)\to \forall y(E(y, f_\psi(\vec x, w))\to A(\vec a, f_\psi(\vec x, w)))\to B(\vec{\red{b}}, w))$$
with $f_\psi$ being a new function symbol.

@@

Even if we \sout{don't}\red{don't} have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$g_{\psi_1}(\vec m_1, g_{\psi_2}(\vec m_2, \dots g_{\psi_n}(\vec m_n, s)\dots))$$ is sufficient.
In particular, this allows us to completely eliminate $\preceq$, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and adjust DomainSubset$(u, w)$ and Persistent$(u, w)$, e.g. remove quantification over $w$ and the precondition $u\preceq w$ from $K(\varphi)$ and replace DomainSubset$(u, w)$ with
$$
	\bigwedge\{\forall u\forall x\forall\vec y(E(x, u)\to E(x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\}
$$
and Persistent$(u, w)$ with
$$
	\bigwedge\{\forall\vec x\forall\vec y\forall u(A(\vec x, u)\to A(\vec x, f_\psi(\vec y, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q, \text{$A$ a predicate in $\varphi$}\}.
$$Putting all of this together yields the following translation:
\begin{theorem}\label{thm:fo-translation}
	Let $\mathcal R,\mathcal X,\mathcal Q$ be as above. For each $n$-ary predicate $A$ define a new $n+1$-ary predicate of the same name. For every $\psi\in\mathcal \mathcal X\cup\mathcal Q$ define a new function symbol $f_\psi$. Consider a new binary predicate $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal R^\#$ by including
	$$
		\forall u\forall \vec x\left(\vec E(\vec x, u)\to\bigwedge_i A_i(\vec a_i, u)\to\bigvee_j B_j(\vec b_j, u)\right)
	$$
 	for each flat clause
 	$
 		\forall \vec x\left(\bigwedge_i A_i(\vec a_i)\to\bigvee_j B_j(\vec b_j)\right)\in\mathcal R,
 	$
 	$\mathcal X^\#$ by including
 	$$
 		\forall u\forall \vec x\left(A(\vec a, f_\psi(\vec x, u))\to B(\vec b, f_\psi(\vec x, u))\to C(\vec c, u)\right)
 	$$
 	for each implication clause
	$
		\psi = \forall \vec x\left(\left(A(\vec a)\to B(\vec b)\right)\to C(\vec c)\right)\in\mathcal X,
	$
	and $\mathcal Q^\#$ by including
	$$
		\forall u\forall \vec x\left(\forall y\left(A(\vec a, f_\psi(\vec x, u))\right)\to B(\vec b, u)\right)
	$$
	for each implication clause
	$
		\psi = \forall \vec x\left(\forall y\left(A(\vec a)\right)\to B(\vec b)\right)\in\mathcal X.
	$
	Then $\varphi$ is intuitionistically valid if and only if
	$$
		\varphi^\# := K(\varphi)\to\text{World}(s)\to\vec E(\vec r, s) \to \left(\bigwedge\mathcal R^\#\wedge\bigwedge\mathcal X^\#\wedge\bigwedge\mathcal Q^\#\right)\to P^\#(s)
	$$
	is classically valid, where $s, \vec r$ are new constants introduced by Herbrandization of free variables,
	\begin{align*}
		K(\varphi) := &\:\bigwedge\left\{\forall\vec z\forall x\forall u((E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal X\cup\mathcal Q\right\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ a function in $\varphi$}\}\right)\wedge\\
		&\resizebox{.9\hsize}{!}{$\:\bigwedge\left\{\forall\vec z_1\forall \vec z_2\forall u(A(\vec z_1, u)\to A(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal X\cup\mathcal Q$, $A$ a predicate in $\varphi$}\right\}$.}
	\end{align*}
	The size of the obtained formula is linear in the input. However $|\mathcal X| + |\mathcal Q| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary predicate has been extended to a $n+1$-ary predicate.
\end{theorem}

@@

We have benchmarked our first-order translation on version 1.1.2 of the ILTP problem set~\cite{iltp} by performing our translation and then examining the generated classical problem using Vampire~\cite{kovacs2013first}.
For comparison the state-of-the-art intuitionistic provers ileanCoP 1.2~\cite{otten2008leancop} and nanoCoP-i 2.0~\cite{otten2021nanocop} were used. All benchmarks were conducted on an Intel Core i5-8400 2.80GHz machine running Linux with a time limit of 10s per problem. The problem \textsc{SYN007+1} was removed as our normal form procedure would explode the size of \red{the} instance.

@@

In general, our approach falls short of the state of the art. 
However when grouping the benchmarks by $|\mathcal Q|$ and $|\mathcal X|$, see Figure~\ref{fig:b},
we observe good performance of our approach for $|\mathcal Q| = 0$ and $|\mathcal X| \leq 2$\sout{resp. $|\mathcal X| \geq 6$}.
